---
title: "Assignment 4"
author: "Laura Ingulsrud"
date: "11/15/2018"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo = FALSE, message = FALSE}

# Load necessary packages
library(tidyverse)
library(kableExtra)
library(vcdExtra)
library(magrittr)
library(effsize)
library(pwr)
library(knitr)
library(ggplot2)
library(ggsignif)

```

```{r,echo = FALSE, message = FALSE}

# Read in .csv files
lob_size_abundance <- read.csv("lobster_size_abundance.csv")
lob_traps<- read.csv("lobster_traps.csv")

```

```{r,echo = FALSE, message = FALSE}

# Problem 1. Lobster abundance and fishing pressure (2012 - 2017). Describe trends in lobster abundance (counts) and fishing pressure (trap buoys) at the five locations from 2012 - 2017. Ignore transect information - we are only interested in evaluating abundance and pressure on the order of SITE. Note: you are not expected to use regression here - just think of ways to clearly describe annual totals visually and in text, noting important trends, events, and differences.


lob_abundance <- lob_size_abundance %>% # Create simple data frame of lobster abundance from lob_size_abundance data frame
  select(YEAR, SITE, COUNT) %>% # Select only YEAR, SITE, and COUNT columns
  filter(COUNT!=0) %>%  # Filter out all values of 0 in COUNT column
  rename(Site= SITE) %>% 
  mutate(Site= case_when(Site== "CARP"~"Carpinteria",
                         Site== "NAPL"~ "Naples",
                         Site== "MOHK"~"Mohawk",
                         Site== "AQUE"~"Arroyo Quemado",
                         Site== "IVEE"~"Isla Vista"
                      ))


lob_traps_simple <- lob_traps %>% # Create simple data frame of lobster traps from lob_traps data frame
  select(YEAR, SITE, TRAPS) %>% # Select only YEAR, SITE, and TRAPS columns
  filter(SITE!="AHND",
         SITE!="ABUR",
         SITE!="AHND to AQUE",
         SITE!="GOLB") %>%  # Filter out all sites except the 5 LTER locations: AQUE, NAPL, MOHK, IVEE, and CARP
  rename(Site= SITE) %>% 
    mutate(Site= case_when(Site== "CARP"~"Carpinteria",
                         Site== "MOHK"~"Mohawk",
                         Site== "AQUE"~"Arroyo Quemado" ))



# Make scatter plot for annual lobster abundance by site (2012-2017)
abundance <- lob_abundance %>% 
 group_by(Site,YEAR) %>% 
 summarize(total_count=round(sum(COUNT)), digits=1)%>% 
 ggplot(aes(x=YEAR,y= total_count)) +
 geom_point()+
 facet_wrap(~Site, scale="free") +
 xlab("Year") +
 ylab("Total Count")
   
abundance # Call "abundance" scatter plot

# Make a column graph of total lobster abundance by site
abundance_col <- ggplot(lob_abundance, aes(x=YEAR, y=COUNT))+
  geom_col()+
  facet_wrap(~Site, scale="free")

abundance_col # Call "abundance_col" column graph

# Make a line graph of total lobster abundance within each site (2012-2017)
abundance_line <- as.data.frame(lob_abundance) %>% 
  expand.dft(freq = "COUNT")%>% 
  count(YEAR, Site) %>% 
  ggplot(aes(x=YEAR, y=n))+
  geom_line(aes(color=Site))+
  xlab("Year")+
  ylab("Lobster Abundance (Count)")+
  theme(text = element_text(family = "Times New Roman")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  geom_point(aes(color=Site))


abundance_line # Call "abundance_line" line graph

# Make scatter plot for annual traps set by site (2012-2017)
traps <- lob_traps_simple %>% 
 group_by(Site,YEAR) %>% 
 summarize(total_count=round(sum(TRAPS)), digits=1)%>% 
 ggplot(aes(x=YEAR,y= total_count))+
 geom_point()+
 facet_wrap(~Site, scale="free")+
 xlab("Year")+
 ylab("Total Traps")+
scale_y_continuous(expand = c(0,0)) 
   
traps # Call "traps" scatter plot

# Make a column graph of total traps set by site 
traps_col<- ggplot(lob_traps_simple, aes(x=YEAR, y=TRAPS))+
  geom_col()+
  facet_wrap(~Site, scale="free")

traps_col # Call "traps_col" column graph

# Make a line graph of total traps at each site (2012-2017)
traps_line<- as.data.frame(lob_traps_simple) %>% 
  expand.dft(freq = "TRAPS") %>% 
  count(YEAR, Site) %>% 
  ggplot(aes(x=YEAR, y=n))+
  geom_line(aes(color= Site))+
  xlab("Year")+
  ylab("Fishing Pressure (Traps)")+
  theme(text = element_text(family = "Times New Roman")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  geom_point(aes(color=Site))
 
traps_line # Call "traps_line" line graph

```
Figure 1. 

Abundance of Lobster from 2012-2017:
AQUE stays relatively consistent
CARP starts increasing after 2015, with a big spike between 2016 and 2017
IVEE increases from 2012 to 2015, decreases from 2015 and 2016, and increases again to 2017
MOHK spikes between 2014 and 2015, then steadily decreases to 2017
NAPL increases steadily between 2012 and 2015, then decreases steadily to 2017
IVEE, MOHK, and NAPL all increase in abundance between 2013 and 2015, but then all decrease to 2016. Out of these 3 sites, only IVEE rapidly increases in abundance from 2016 to 2017; the other 2 sites (MOHK & NAPL) decrease slightly.
CARP and AQUE both don't have large changes in abundance from 2012 to 2015; after 2015, AQUE abundance decreases a little, and CARP increases, with a large increase from 2016 to 2017.

Fishing Pressure from 2012-2017 at Non-MPA Sites:

```{r,echo = FALSE, message = FALSE}

# Problem 2. Compare mean lobster size by site in 2017. Compare mean lobster sizes (carapace length (mm)) across the five sites for lobster observations collected in 2017. Warning: the size data are not in tidy format. There are rows that contain size information for multiple lobsters observed (e.g., if the researcher saw 3 lobsters all with carapace length ~ 60 mm, then they will have a single row where COUNT = 3 and SIZE = 60). You’ll want to get this into case format - where each lobster has its own row - before doing statistical analyses. There are many ways to do this. One hint: function expand.dft in the vcdExtra package (it doesn’t like tibbles, so you might need to coerce to data.frame first). *Anova boxplot and error bars


# put lobster size data into case format
lob_abun_simple<-expand.dft(lob_size_abundance, freq="COUNT")

#sumarize abundance data
lob_simple<-lob_abun_simple %>% 
  filter(YEAR == 2017) %>% 
  select(SITE, SIZE) %>% 
  group_by(SITE) %>% 
  summarize(Mean= mean(SIZE),
           SD= sd(SIZE),
           Median= median(SIZE),
           Maximum= max(SIZE),
           SampleSize= length(SIZE),
           SE= SD/sqrt(SampleSize))
           
lob_simple

# make column graph of mean lobster sizes 
lob_col<-ggplot(lob_simple, aes(x = SITE, y = Mean)) +
  geom_col(aes(fill= "pink"), color="white", width = 0.6) +
  geom_errorbar(aes(ymin= Mean-SD, ymax= Mean+SD, width=.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,110)) +
  xlab("Site")+
  ylab("Mean Lobster Size (mm)") +
  scale_x_discrete(labels = c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples")) +
  geom_signif(y_position = c(93,100), xmin = c(2,3), xmax = c(5,5), annotation = c("p = 0.023", "p = 0.004"), tip_length = 0.1, size = 0.5, textsize = 3) +
  theme(text = element_text(family = "Times New Roman")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        legend.position= "none",
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
  
lob_col

#filter dataset for just 2017 values
Lob_2017_simple <- lob_abun_simple %>% 
  filter(YEAR == 2017)

#Anova of 2017 Lobster size Data across all 5 sites
anov_lobby <- aov(SIZE ~ SITE, data = Lob_2017_simple)
anov_sum <- summary(anov_lobby)

anov_sum

lobby_boxplot <- ggplot(Lob_2017_simple, aes(x= SITE, y= SIZE)) +
  geom_boxplot(aes(fill = SITE, colour = SITE), alpha = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar") +
  theme_classic() +
  labs(x= "Site", y= "Lobster Size" ) +
  theme(legend.position = "NONE")

lobby_boxplot

# At least two samples were taken from populations with different means. Which ones are different? All three are different from each other? Or something else?

# Post-hoc testing using Tukey's HSD

lob_Tukey <- TukeyHSD(anov_lobby)
lob_Tukey
# Naples Reef and Isla Vista differ significantly

```

Mean lobster sizes for Naples Reef and Isla Vista and Naples and Carpinteria are significantly different (F(df) = 4.78, p = 0.004, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).


```{r,echo = FALSE, message = FALSE}

# 3. Changes in lobster size at MPA and non-MPA sites (comparing only 2012 and 2017 sizes). From the data description (http://sbc.lternet.edu/cgi-bin/showDataset.cgi?docid=knb-lter-sbc.77): “Data on abundance, size and fishing pressure of California spiny lobster (Panulirus interruptus) are collected along the mainland coast of the Santa Barbara Channel. Spiny lobsters are an important predator in giant kelp forests off southern California. Two SBC LTER study reefs are located in or near the California Fish and Game Network of Marine Protected Areas (MPA), Naples and Isla Vista, both established as MPAs on 2012-01-01. MPAs provide a unique opportunity to investigate the effects of fishing on kelp forest community dynamics. Sampling began in 2012 and is ongoing.” At Isla Vista and Naples Reef, the two protected MPA sites (with zero fishing pressure), how do lobster sizes in 2012 and 2017 compare? At the non-MPA sites?

MPA12 <-lob_abun_simple  %>% 
  filter(YEAR == "2012" | YEAR == "2017") %>% 
  mutate(PROTECTED = case_when(
    SITE == "IVEE" ~ "MPA",
    SITE == "NAPL" ~ "MPA",
    SITE == "AQUE" ~ "Non-MPA",
    SITE == "CARP" ~ "Non-MPA",
    SITE == "MOHK" ~ "Non-MPA")) %>% 
  mutate(SITE= case_when(SITE== "CARP"~"Carpinteria",
                         SITE== "NAPL"~ "Naples",
                         SITE== "MOHK"~"Mohawk",
                         SITE== "AQUE"~"Arroyo Quemado",
                         SITE== "IVEE"~"Isla Vista"
                      )) 
  

MPA_Graph_12<- ggplot(MPA12, aes(x= SITE, y= SIZE))+
  geom_boxplot(aes(fill = PROTECTED))+
  facet_wrap(~ YEAR)+
  theme_classic()+
  theme(text = element_text(family = "Times New Roman")) +
  scale_x_discrete(limits = c("Arroyo Quemado", "Carpinteria", "Mohawk", "Isla Vista", "Naples"))+
  ylab("Carapace Length (mm)")+
  xlab("Site")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))


MPA_Graph_12 + theme(panel.spacing = unit(3, "lines"))




```



```{r,echo = FALSE, message = FALSE}

#Perform 5 two sample t-tests to test if lobster sizes differ significantly between 2012 and 2016 (one t-test per site)

#H0: means are equal (differnce in means = 0)
#Ha: means are NOT equal (difference in means is NOT 0)

# Simplify data for NAPL
NAPL_abun_12<- lob_abun_simple %>% 
  filter(SITE=="NAPL", YEAR== "2012") %>% 
  pull("SIZE")
NAPL_abun_12

NAPL_abun_17<- lob_abun_simple %>% 
  filter(SITE=="NAPL", YEAR== "2017") %>% 
  pull("SIZE")
NAPL_abun_17


#f-test to test if variances are equal
NAPL_ftest<-var.test(NAPL_abun_17, NAPL_abun_12)
NAPL_ftest
#p-value =  0.7685 thus variances are equal

#two sample, two sided t-test
t_test<- t.test(NAPL_abun_17, NAPL_abun_12, var.equal=TRUE)
t_test 
#p-value =0.5002
#thus, there IS significant diference between the two means

#Two sided t test
t_test<- t.test(NAPL_abun_17, NAPL_abun_12, var.equal=TRUE, alternative = "greater")
t_test 
#p-value = 0.2501


#test effect size
NAPL_effect_size<- cohen.d(NAPL_abun_17, NAPL_abun_12)
NAPL_effect_size  
# cohens d = 0.2834216 (small)
# Thus, we have a small effect size

```

```{r,echo = FALSE, message = FALSE}


# maybe a man whiney u?

# Simplify data for IVEE
IVEE_abun_12<- lob_abun_simple %>% 
  filter(SITE=="IVEE", YEAR== "2012") %>% 
  pull("SIZE")
IVEE_abun_12

IVEE_abun_17<- lob_abun_simple %>% 
  filter(SITE=="IVEE", YEAR== "2017") %>% 
  pull("SIZE")
IVEE_abun_17

#f test to test if variances are equal
IVEE_ftest<-var.test(IVEE_abun_17, IVEE_abun_12)
IVEE_ftest
#p-value 0.307 thus variances are equal

#two sample, two sided t-test
t_test<- t.test(IVEE_abun_12, IVEE_abun_17, var.equal=TRUE)
t_test 
#p-value = 0.0599
#thus, there IS a significant diference between the two means

t_test<- t.test(IVEE_abun_12, IVEE_abun_17, var.equal=TRUE, alternative= "greater")
t_test 
#p-value = 0.9701

#test effect size
IVEE_effect_size<- cohen.d(IVEE_abun_17, IVEE_abun_12)
IVEE_effect_size  
# cohens d = 0.3775177 (small)
# Thus, we have a small effect size

```

```{r,echo = FALSE, message = FALSE}

# Simplify data forAQUE
AQUE_abun_12<- lob_abun_simple %>% 
  filter(SITE=="AQUE", YEAR== "2012", SIZE!= "-99999") %>% 
  pull("SIZE")
AQUE_abun_12

AQUE_abun_17<- lob_abun_simple %>% 
  filter(SITE=="AQUE", YEAR== "2017") %>% 
  pull("SIZE")
AQUE_abun_17

#f test to test if variances are equal
AQUE_ftest<-var.test(AQUE_abun_17, AQUE_abun_12)
AQUE_ftest
#p-value = 0.2986 thus variances are  equal

#two sample, two sided t-test
t_test<- t.test(AQUE_abun_17, AQUE_abun_12, var.equal=TRUE)
t_test 
# p-value = 0.2097
#thus, there IS a significant diference between the two means

t_test<- t.test(AQUE_abun_17, AQUE_abun_12, var.equal=TRUE, Alternative="Greater")
t_test 

#p-value = 0.2097

#test effect size
AQUE_effect_size<- cohen.d(AQUE_abun_17, AQUE_abun_12)
AQUE_effect_size  
# cohens d = 0.2563169 (small)
# Thus, we have a small effect size

```

```{r,echo = FALSE, message = FALSE}

# Simplify data for CARP
CARP_abun_12<- lob_abun_simple %>% 
  filter(SITE=="CARP", YEAR== "2012") %>% 
  pull("SIZE")
CARP_abun_12

CARP_abun_17<- lob_abun_simple %>% 
  filter(SITE=="CARP", YEAR== "2017") %>% 
  pull("SIZE")
CARP_abun_17

#f test to test if variances are equal
CARP_ftest<-var.test(CARP_abun_17, CARP_abun_12)
CARP_ftest
#p-value=  0.2043 thus variances not equal


#two sample, two sided t-test
t_test<- t.test(CARP_abun_17, CARP_abun_12, var.equal=TRUE)
t_test 
#p-value = 0.1819
#thus there is a significant difference between the two means

t_test<- t.test(CARP_abun_17, CARP_abun_12, var.equal=TRUE, alternative = "less")
t_test 

# p-value = 0.09095

# test effect size
CARP_effect_size<- cohen.d(CARP_abun_17, CARP_abun_12)
CARP_effect_size  
#cohens d =  -0.1594364 (small)    
#the cohens d values turned up NEGATIVE so I'll preform a power test to find a more accureate Cohens D
CARP_power<- pwr.t.test(n = 783, d = NULL, sig.level = 0.05, power = 0.8)
CARP_power
#using the power test I found that the cohens d value is 0.1416723 which means there is a small effec size

```

```{r,echo = FALSE, message = FALSE}

#Simplify data for MOHK
MOHK_abun_12<- lob_abun_simple %>% 
  filter(SITE=="MOHK", YEAR== "2012") %>% 
  pull("SIZE")
MOHK_abun_12

MOHK_abun_17<- lob_abun_simple %>% 
  filter(SITE=="MOHK", YEAR== "2017") %>% 
  pull("SIZE")
MOHK_abun_17

#f test to test if variances are equal
MOHK_ftest<-var.test(MOHK_abun_17, MOHK_abun_12)
MOHK_ftest
#p-value = 0.1509 thus variances are equal 
# thus there is not a significant difference between the two means


#two sample, two sided t test
t_test<- t.test(MOHK_abun_17, MOHK_abun_12, var.equal=TRUE)
t_test 
#p value > .0001
#there is NOT a significant difference between the two means

# cohens-d test
MOHK_effect_size<- cohen.d(MOHK_abun_17, MOHK_abun_12)
MOHK_effect_size  
#cohens d =  -0.5408116 (medium)    
#the cohens d values turned up NEGATIVE so I'll preform a power test to find Cohens D
MOHK_power<- pwr.t.test(n = 261, d = NULL, sig.level = 0.05, power = 0.8)
MOHK_power
#using the power test I found that the cohens d value is  0.2457094 which means there is a small effect size
```

```{r}

# Problem 4. Proportions of “legal” lobsters at the 5 sites in 2017. The legal minimum carapace size for lobster is 82.6 mm. What proportion of observed lobsters at each site are above the legal minimum? Does that proportion differ significantly across the 5 sites? Note: We’ll be doing chi-square in labs next week, or go ahead with maximum resourcefulness and check out the chisq.test() function on your own!

# Make a contingency table for lobster carapace size:
lob_size <- lob_abun_simple %>% # Create data frame for lobster carapace size from "lob_abun_simple" data frame (created in Problem 2), which has data in tidy format
  filter(YEAR == 2017) %>% 
  mutate(Legality = ifelse(SIZE < 82.6, "ILLEGAL", "LEGAL")) %>% 
  select(SITE, Legality) %>% 
  count(SITE, Legality) %>% 
  spread(Legality, n) %>% 
  select(-SITE)
  
rownames(lob_size) <- c("AQUE", "CARP", "IVEE", "MOHK", "NAPL")

lob_size

lob_prop <- prop.table(as.matrix(lob_size), 1) # Look at the actual proportions
lob_prop

# Perform chi-square test for independence on the count data (not proportions)
lob_x2 <- chisq.test(lob_size)
lob_x2

lob_x2$stdres # If beyond 2, pay attention for significance...
# There IS a significant association between legal lobster size between the 5 sites! Look at 'Values' section to see what things you can call from the output.

# create kable table of proportion of legal lobsters at each site
lob_table <- kable(lob_prop) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = F, position = "left")
lob_table
# need to change to percentages in the table

```

The proportion of observed lobsters at a legal size differed significantly across the 5 sites ($\chi^2$(`r lob_x2$parameter`) = `r round(lob_x2$statistic, 2)`, p < 0.001, $\alpha$ = 0.05.) Most notably, Mohawk Reef had a higher proportion of illegal lobster size, while Naples had a higher proportion of legal lobster size compared to the rest of the 5 sites.
